% Clever thing to minimize the files I need to edit to change how the files are
% looking. The following one line comment is to trick VIM-LaTeX
% \documentclass{article}
\input{./../template_preamble.tex}

% Custom packages

\title{Labels and Cross-referencing in \LaTeX{}}
\author{Ignas Anikevicius}

\begin{document}

\maketitle

This will be a short but \emph{very} meaningful tutorial on cross-references in
\LaTeX{}
%
What makes \LaTeX{} cross-referencing so good is that you can literally refer to
    almost any object you want, which makes it very useful.
%
The fact, that this task is trivial to do makes it very difficult to ignore it.

%    
Everything works via \verb|label| command, which defines a \emph{makrker}, to
    which you can refer later on via \verb|ref| command.
%
So for example, referencing to a section called ``Introduction'' you just have
    to specify it with the following code:
%
\begin{lstlisting}
\section{Introduction}
\label{sec:intro}
\end{lstlisting}
%
or
%
\begin{lstlisting}
\section{Introduction \label{sec:intro}}
\end{lstlisting}
%
Both code variants will work and when you would execute a command
    \verb|\ref{sec:intro}|, you would get a section number.
%
For getting a page-number you would similarly execute 
    \verb|\pageref{sec:intro}| command.

%
Now note the prefix I have given to the newly created label.
%
Since one can refer to almost any object, it might get slightly confusing once
    you have more than 10 references.
%
In order to avoid this, most \LaTeX{} users make labels with various prefixes,
    which denote what type of object it is.
%
Common prefixes are as follows:
%
\begin{itemize}
    \item \emph{sec:} for sections
    \item \emph{subsec:} for sections or you can use \emph{ssec:}
    \item \emph{eq:} for equations
    \item \emph{fig:} for figures
    \item \emph{tab:} for tables
\end{itemize}

Note, that the list of the prefixes is completely trivial, so you can use
    whatever prefixes you like.
%
If you use packages, which provide new commands, or environments, then it would
    be good idea to come up with prefixes for these things as well.
%
\begin{itemize}
    \item \emph{schem:} for schemes
    \item \emph{lst:} for code listings
    \item \emph{cheq:} for chemical reaction equations
\end{itemize}

\emph{NOTE}, the prefix convention is not obligatory, so one can not use it, but
    in the end, for bigger documents, it makes reading, and writing the document
    so much easier.
%
Therefore, it would be a good idea to use it \emph{always}.

For more information on cross-referencing, please read a chapter on
    \uurl{https://secure.wikimedia.org/wikibooks/en/wiki/LaTeX/Labels_and_Cross-referencing}{Labels
    and Cross-referencing} on the 
    \uurl{https://secure.wikimedia.org/wikibooks/en/wiki/LaTeX}{\LaTeX{} wikibook}.


\end{document}

% Editor configuration:
% vim: tw=80:spell:spelllang=en_gb


