% Clever thing to minimize the files I need to edit to change how the files are
% looking. The following one line comment is to trick VIM-LaTeX
% \documentclass{article}
\input{./../template_preamble.tex}

% Custom packages

% Version of the XeTeX logo that doesn't depend on Ǝ being available in the font
% NB: not suitable for use in italic!
% Requires XeTeX 0.7 or later
\def\reflect#1{{\setbox0=\hbox{#1}\rlap{\kern0.5\wd0
  \special{x:gsave}\special{x:scale -1 1}}\box0 \special{x:grestore}}}
\def\XeTeX{\leavevmode
  \setbox0=\hbox{X\lower.5ex\hbox{\kern-.15em\reflect{E}}\kern-.1667em \TeX}%
  \dp0=0pt\ht0=0pt\box0 }

\title{Fonts in \LaTeX{}}
\author{Ignas Anikevicius}

\begin{document}

\maketitle

\begin{abstract}
    Being able to customize fonts in your \verb|.tex| document can be tricky at
    times, or as easy as loading a package or two. In this tutorial I will talk
    through various ways on how to accomplish it.
\end{abstract}

\tableofcontents

\section{The Font Catalogue}

There is a very good resource for fonts, which outlines how to make different
fonts working and it includes notes whether the mathematics mode is supported
with those fonts. The resource is located at
\href{http://www.tug.dk/FontCatalogue/}{\LaTeX\ font catalogue} website and you
will find the information you search.

As for fonts without the support for mathematical symbols, I would advice you
\textbf{against} using them unless you do not use any mathematical  typesetting
in your documents (e.g. this is good for letters and other text-only documents).
However, if you need to do that, please consult professionals, some geek-written
blogs or just people who are more experienced in that field.

Now let's talk about commands involved setting different fonts. Usually you
should declare in your preamble, that you will use Type1 fonts, which look
better and most of the scientific journals require them. Here is how you do it:
\begin{lstlisting}
\usepackage[T1]{fontenc}
\end{lstlisting}

And then you put a package responsible for defining the font:
\begin{description}
    \item[times fonts] Just put \verb|\usepackage{mathptmx}| after the previous
        command;
    \item[palatino] Alternatively for palatino fonts, just write
        \verb|\usepackage{mathpazo}| command;
\end{description}

\section{UTF8 usage}

\section{\XeTeX{} usage}

\end{document}

% Editor configuration:
% vim: tw=80:spell:spelllang=en_gb


