% Clever thing to minimize the files I need to edit to change how the files are
% looking. The following one line comment is to trick VIM-LaTeX
% \documentclass{article}
\input{./../template_preamble.tex}

% Custom packages

\title{How to use GIT version control system to backup and work collaboratively
on the same documents}
\author{Ignas Anikevicius}

\begin{document}

\maketitle

\LaTeX{} is a very good tool for this purpose as the files are created in ASCII
format and everybody can read it the same way.

\tableofcontents

\section{What is version control system (VCS)?}

There are many models of VCS, but the main idea in all of them remains the same.
This is the ability of registering the changes to the files. A typical workflow
can be described as follows:
\begin{enumerate}
    \item Repository is initialized with initial files.
    \item Then a person copies the repository
    \item makes some changes to the files
    \item submits them back to the repository 
    \item clever algorithms detect  which part(s) of the files were altered 
    \item The files are updated and the changes being made are saved
    \item Go back to point 2.
\end{enumerate}

As you see from this description it is clear that the whole history of how the
files were changing is saved and, thus, they can be restored to any previous
state. In addition to this being such a good back up tool, there are several
other advantages one should be aware of:
\begin{itemize}
    \item It takes much less space than having multiple folders with different
        versions of the files
    \item One can spot what was changed much more easily
    \item People can work on different parts of the file at the same time.
\end{itemize}

The last feature is the most useful for \LaTeX{} users once they need to work
together with someone.

\section{Using GIT and \LaTeX{} together}

\subsection{Introduction on Git}

Git is a fully open source Distributed Version Control System (DVCS) used for
many major Open Source Software (OSS) projects.
%
This means, that you get an advanced tool totally for free and you have got an
army of community members around the world from whom you can learn about git and
find new 'tricks' almost every day.
%
Distributed VCS from a simple VCS differs in one aspect which makes the former
much more flexible and simply better than the latter - every user has its own
copy of the entire repository and can work on it in any possible way.
%
This has some implications:
\begin{itemize}
    \item The user can work off-line and then when he goes back on-line he can
        submit the number of changes to the main repository.
    \item If the main repository goes down or is compromised, you can get the
        source code back to its original state simply because every user has a
        copy of it.
\end{itemize}



What is more, git has introduced 'branching' mechanism into VCS.
%
Branching means, that you can create a branch of a repository and then make some
changes on it and then merge it back whenever you want.
%
Or in the same way you can just abandon it, if you think that the changes you
made where simply a mistake and you want the previous state of the repository
back.
%
Or you can even 'cherry-pick' some changes from the branch and merge it back to
the 'master' branch and discard the others.


%TODO
\textcolor{Red}{\bfseries 
    Tell more about Git Pull Push requests.
}

This way one can have a very powerful tool to manage all the changes to the
files on the system.

\subsection{Using Git}

Git system is very easy to use if you are used to working in shell (such as Bash
or Zsh).
%
However, it might require more adjustment for those who have always been using
Graphic User Interface and are not as comfortable without any buttons or icons
to press.
%
Luckily there are GUI software, which might help you interface with a Git
repository very easily:
\begin{itemize}
    \item Some
    \item Or
    \item Links
    \item on
    \item GIT
    \item GUIS
\end{itemize}

\subsection{Combining \LaTeX{} and Git}

\subsection{Some thoughts on more sophisticated approaches}

\section{Using other VCS solutions}

Using other VCS solutions is possible and highly recommended for people, who
find GIT too hard. 
%
One very good alternative might be the Mercurial versioning system, which
receives a lot of praises amongst its users.
%
However, Mercurial seems to be not as popular as Git and hence, the resources on
the web might not be as elaborate.

For those people who know Subversion and CVS and claim that they are really good
alternatives, I would advise to look into Mercurial and Git very seriously and
consider switching over.
%
It is because the old SVN and CVS systems are slower, not as space-efficient and
they are not as flexible.

However, there is a SVN repository on the Chemistry Department servers, which
would help you very much in setting up VCS repository which would not be public.
%
If you do not mind spending time learning an obsolete technology, then please
start using SVN as soon as possible.
%
There are already good guides about how to use Subversion with \LaTeX{} and you
can find them on the
\uurl{https://secure.wikimedia.org/wikibooks/en/wiki/LaTeX/Collaborative_Writing_of_LaTeX_Documents\#The_Version_Control_System_Subversion}{\LaTeX{}
wikibook}.
%
Otherwise please ask the Computer Officers in the Department about how the
things are going towards a git repository on their servers.

\end{document}

% Editor configuration:
% vim: tw=80:spell:spelllang=en_gb


