%        File: questions.tex
%     Created: Wed Jul 06 02:00  2011 B
% Last Change: Wed Jul 06 02:00  2011 B
%
\documentclass[a4paper]{article}
\usepackage{fullpage}

\begin{document}


\title{Questionnaire on LaTeX website}

\author{Ignas Anikevicius\\ email: ia277@cam.ac.uk}

\maketitle

\thispagestyle{empty}

Hello, I am Ignas Anikevicius and I am currently working under Dr. Scherman to
produce a website that would serve all members of the department from Part II
students to group leaders containing all things that are Chemistry and \LaTeX
related.

I was wondering if you could answer some questions for me, which would
significantly benefit the quality and the content of the website.

\paragraph{The preliminary list of the content includes:}

\begin{itemize}
    \item notes, links, templates, instructions and downloadable .sty and .cls
        files to produce a Cambridge Department of Chemistry PhD thesis, CPGS
        report and part III project report (NB here I would really like to get
        input and hopefully have several examples which can be used for
        interested students)
    \item notes, links, templates, instructions and downloadable .sty and .cls
        files to produce manuscripts for several scientific journals (ACS, NPG,
        Willey-VHC and possibly others)
    \item notes, links, templates on how to produce a CV in LaTeX 
    \item notes, links on how to properly set up a computer to start using LaTeX
    \item any other requests or ideas?
\end{itemize}

\paragraph{The questions:}
\begin{enumerate}
    \item For what purpose do you usually use \LaTeX?
    \item Do you see any real drawbacks in using \LaTeX for your all
        word-processing needs?
    \item In addition to the mentioned content of the website, what else would
        you like to see there?
    \item Could you give names of scientific journals to which you have
        submitted papers using \LaTeX? For which journals you had to use a
        different typesetting system (such as Word or Open Office)?
    \item Could you share any .tex, .cls and .sty files which you had write on
        your own? This would serve as a great example for templates and website
        making in general.
    \item Would you like to use \LaTeX  for typing documents in other languages
        than English?
    \item Would you like to add anything?
\end{enumerate}

\end{document}

% vim: tw=80
