\input{./preamble_cv.tex}

\begin{document}

\makeheading{Ignas Anikevicius}

\section{Contact Information}
%
% NOTE: Mind where the & separators and \\ breaks are in the following
%       table.
%
% ALSO: \rcollength is the width of the right column of the table
%       (adjust it to your liking; default is 1.85in).
%
\newlength{\rcollength}\setlength{\rcollength}{1.95in}%
%
\begin{tabular}[t]{@{}p{\textwidth-\rcollength-.3in}p{\rcollength}}
  St. Catharine's College       & \\
  Trumpington Street            & \hfill Mobile: 07535170591 \\
  Cambridge, CB2 1RL            & \hfill E-mail: anikevicius@gmail.com \\
  United Kingdom                & \hfill {\bf or} ia277@cam.ac.uk \\
\end{tabular}

\section{Citizenship}
Lithuanian, EU citizen

\section{Education}
\textbf{University of Cambridge, St. Catharine's College
\hfill 2010 to present}\\
1\textsuperscript{st} year Natural Sciences Tripos

\begin{innerlist}
  \item Physics, supervised by Dr. Stephen Morris
  \item Mathematics, supervised by the {\em Director of Studies}, Dr. Sergei N.
          Taraskin (snt1000@cam.ac.uk)
  \item Chemistry, supervised by Dr. Peter Wothers
  \item Materials Science, supervised by Dr. John Little
\end{innerlist}

\textbf{Vilnius lyceum, International Baccalaureate
\hfill 2006-2009}

\begin{innerlist}
  \item Mathematics HL, Chemistry HL, Physics HL, English B HL
  \item Psychology SL, Lithuanian SL
  \item Extended Essay from Chemistry
\end{innerlist}

\section{Honours \& Awards}
\textbf{Cambridge}:
\begin{innerlist}
  \item Thompson award from St. Catharine's College given for academic
    excellence.
  \item Work through \emph{Michaelmas} and \emph{Lent} Terms was assessed as
    1\textsuperscript{st} class in every subject.
\end{innerlist}

\textbf{International Chemistry Olympiads (IChO):}
\begin{innerlist}
  \item 41th IChO in Cambridge - Silver Medal
    \hfill July, 2009
  \item 40th IChO in Budapest - Bronze Medal
    \hfill July, 2008
\end{innerlist}

\section{Languages}
\begin{loneinnerlist}
  \item Lithuanian - native speaker
  \item English - fluent 
  \item French - intermediate (B1)
  \item Russian - basics
\end{loneinnerlist}

\section{Computing Skills}
\begin{loneinnerlist}
  \item \LaTeX, Python2, Matlab, LUA, basic Bash scripting
  \item Good knowledge of Linux OS (Currently I use Gentoo Linux distribution)
\end{loneinnerlist}

\section{Other Interests \& Activities}

\textbf{Teaching in National Students Academy in Lithuania
\hfill 2010 - present}
\begin{innerlist}
  \item ``Organic Chemistry and Physical Chemistry''
    \hfill January, 2010 
  \item ``Quantum Mechanics in Chemistry'' 
    \hfill May, 2010 \\
    {\em (preparation of students for 42nd IChO)}
  \item ``Equilibrium processes in acid solutions''
    \hfill January, 2011
  \item Distant teaching of the students
\end{innerlist}
\begin{outerlist}
  \item[] \emph{Playing jazz guitar}\\
    I started playing 7 years ago and it never got boring. Jazz music learning
    and analysing the structure of it sometimes keep me busy after a long day.
  \item[] \emph{Photography}\\
    My interest in Physics also lead me to start appreciating
    photography more and it has become one of my hobbies.
\end{outerlist}

\end{document}
% vim: textwidth=80:spell:spelllang=en_gb
