% Clever thing to minimize the files I need to edit to change how the files are
% looking. The following one line comment is to trick VIM-LaTeX
% \documentclass{article}
\input{./../template_preamble.tex}

% Custom packages
\usepackage[]{cite}
\usepackage{framed}

\title{How to make presentations with \LaTeX{}}
\author{Ignas Anikevicius}

\begin{document}

\maketitle

\begin{framed}
\begin{center}
{\color{Red} TODO}
\begin{itemize}
    \item Elaborate more on Beamer package
    \item More on KOMAscript
    \item More on others?
\end{itemize}
\end{center}
\end{framed}

Here I will talk about getting \LaTeX\ working for the production of 
presentations. There are numerous ways how you could do a presentation in
\LaTeX{}. One of
them is to use already existing packages, of which \verb|beamer| is the most
popular one. However, you can use standard document classes as well
\cite{PracTeX-komapres}.

There is also another source of information \cite{latexwikibook:presentations},
which might be a good starting point for using the Beamer document class.

\section{The beamer document class}

As mentioned before, the \verb|beamer| package is the most popular among
scientists and it works quite well with PGF/TikZ packages which might make it
the best solution out there. It is very straight forward to use it --- it
defines several new environments (eg \verb|frame|, \verb|column| etc) and
commands (eg \verb|\frametitle|, \verb|\pause| etc).

The best way to learn it would be to examine the supplied example and comparing
how the source file corresponds to what is produced in the final \verb|.pdf|
document.

\section{Another approach}

This is an alternative method to create presentations. It involves using
\verb|scrartcl| class which is then customized to a great extent. The
decorations of the presentation are done by using the \verb|TikZ| package and
since everything is done from scratch, the user will end up having a unique
theme for his/her presentations.

As you might understand, with a lot of customizability comes slightly steeper
learning curve and one need to spend more time initially to get everything set
up. However, then to convert an article to a presentation is much faster, which
is mainly the point of this method. Also, you do not need to learn knew
commands/environments, which might also be considered as an advantage of this
method.

There is an excellent example and article on the Prac\TeX\
journal.\cite{PracTeX-komapres} Please download the zipped sources where you
will find everything. You can check the presentation .tex file and change
various parameters in the preamble and see how the style of presentation is
affected.

\bibliographystyle{plain}
\bibliography{./../tutorial}

\end{document}

% Editor configuration:
% vim: tw=80:spell:spelllang=en_gb
