% Clever thing to minimize the files I need to edit to change how the files are
% looking. The following one line comment is to trick VIM-LaTeX
\documentclass[
    draft
]{scrartcl}
\input{./../template_preamble.tex}

% Custom things
\usepackage[version=3]{mhchem}
\usepackage[printonlyused]{acronym}
\usepackage[xindy]{glossaries}

\title{Using packages to use dynamic acronyms instead of hard-coding them}
\author{Ignas Anikevicius}

\begin{document}
\maketitle
\tableofcontents
\listoftodos

%
In this document I will try to show you how to create and use a list of acronyms
    in your document the \TeX{} way.
%
One might ask: 'Why you need this? I can type everything manually and it will be
    allright\ldots'
%
I guess that if you write your thesis which is 200 pages long, checking whether
    you have included the definition of the acronym the first time you used a
    phrase is the least you want.

% ----------------------------------------------------------------------
\section{The \pkg{acronym} package}
% ----------------------------------------------------------------------

%
\subsection{An example}

%
\input{acro.tex}

%
\subsection{Code used for the example}

%
\lstinputlisting{acro.tex}

%
As you see, I have defined more acronyms than I have actually used.
%
But, since I have initiated the package with \verb|printonlyused| option as
    follows:
    %
\begin{lstlisting}
\usepackage[printonlyused]{acronym}
\end{lstlisting}
    %
    I do not have to worry about unused acronyms being printed in the list.

%
If by any chance you want some acronyms to be printed out without them using in
    the text, you can achieve it by issuing \verb|\acused{}| command, where in
    the brackets you put the respective label of the acronym.
%
Then the acronym will be printed regardless whether it is mentioned or
    not as the package will think, that it was used.
%
Also, only short version will occur in the text if it will be used after issuing
    the mentioned command.

%
If the \verb|printonlyused| option is omitted, then all defined acronyms will be
    printed out as a result.

% ----------------------------------------------------------------------
\section{The \pkg{glossaries} package}
% ----------------------------------------------------------------------

%
Glossaries package is another possible way to get a similar kind of
    functionality.
%
It has much more functionality and is probably more suited towards people who
    want to create extensive lists and have more than one glossary at a time.
%
With this extensibility comes complexity as it is slightly harder to get the
    glossaries being generated as you need to open command line shell and
    execute some commands.
%
However for simple acronym usage the \pkg{acronym} package is more than enough.
%
But for people who are interested in trying out alternatives, please refer
    to these web-pages and the documentation of the package.

\end{document}

% Editor configuration:
% vim: tw=80:spell:spelllang=en_gb
